---
title: "Starting Words in Octordle"
author: "Dan Lependorf"
date: '2022-02-28'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

I know I know, a million people have already "solved" the best starting word for Wordle. But for that multiplying series of simultaneous Dordle/Quordle/Octordle/Secordle/whatever variants, it seems to me that as the number of puzzles grows, the more important it is to cover the field as broadly as possible, that is, start with a two or even a three word opener that covers all of the 15 most common letters, with no duplicates. Can we figure out what the best three word opener is?

## Grabbing the right list of words

Sure, I could just use any reasonable dictionary, but since this has to be a set of words that works in Octordle, why not go straight to the source? We can grab the full list of words that are allowed in Octordle and use that. I'm going to use the **httr** package to do most of the heavy lifting here.

```{r packages, echo=TRUE, results="hide", message=FALSE, warning=FALSE}
library(tidyverse)
library(httr)
library(knitr)
```

Thankfully, this isn't anything too difficult. If you open Octordle in any browser and take a peek at the source code, you'll find the full list of allowable words right there, in plain text. Let's grab it.

```{r grab_source_code}
octordle_url <- "https://octordle.com/?mode=daily"

octordle_source_code <- octordle_url %>%
    GET() %>%
    content("text")
```

Most of this is a bunch of code that we don't need, but take a look at line 856:

```{r js_example, echo=FALSE, comment=NA}
cat(str_sub(octordle_source_code, start=38892, end=39175))
```

There's a big list of words that's prefaced by `allowed =` and surrounded by quotes. We can use some fairly simple regex to extract that list out from the rest of this.

```{r parse_words}
octordle_words <- octordle_source_code %>%    
    str_match('allowed\\s=\\s"(.*?)"') %>%
    pluck(2) %>%
    str_split(" ") %>%
    pluck(1)

str(octordle_words)
```

In English, that regex essentially translates to: look for anything of the form `allowed = "stuff"` and grab the stuff in between the quotes. That question mark inside of the parentheses capture group tells the regex engine to be lazy about it--in other words, grab the shortest possible match surrounded by quotes, and not to match, say, an ending quotation mark 50 lines later.

The rest of that pipe chain above parses that long list of words into individual words separated by spaces, and there we go, all 10,657 words that are acceptable guesses in Octordle.

## Finding single opener words

Now that we have the list of allowable Octordle words, let's cut that list down to the words that only contain the 15 most common letters in the set. First, we'll have to split each word up into separate letters. I'm also going to take this opportunity to convert this big character vector into a data frame, so that it's easier to work with.

```{r split_words}
split_words <- octordle_words %>%
    tibble(words=.) %>%
    # This map() call inside of mutate() creates a second column of a special form: a list-column. Instead of
    # a column where each entry is a single value, it's a column where each row holds a five-character vector.
    mutate(letters=map(words, ~str_split(.x, "")[[1]]))

split_words
```

And now, grab that column full of letters and count them.

```{r best_letters, comment=NA}
letter_count <- split_words %>%
    unnest(letters) %>%
    count(letters, name="appearances") %>%
    arrange(desc(appearances)) %>%
    filter(row_number() <= 15)

letter_count

best_letters <- letter_count$letters
```

What are the words in the dictionary that contain only those letters and no duplicates?

```{r best_words}
best_words <- split_words %>%
    filter(map_lgl(letters, ~all(.x %in% best_letters)),
           map_int(letters, ~max(table(.x)))==1)

best_words
```

## Finding opener word combinations

We've now cut the word list down from 10,657 down to 2,331. Now that we have a list of words that contain 5 of the 15 most common letters in the acceptable word list, we need to put them together in three word sets that fit together and contain all 15. There are a lot of really inefficient ways of doing this, and I'm not going to assume that this is the fastest way (a triple cross product of 2,331 times 2,331 times 2,331 rows isn't the easiest thing to deal with!), but here's one way that takes about 15 minutes.

```{r the_loop}
opener_combinations <- tibble(word_1=list(),
                              word_2=list(),
                              word_3=list())

# The basic idea here: for each word, remove those letters from the list of best letters, then see which
# words can be constructed out of the remaining letters still available. Then repeat again for the second
# word, to see if any words can be constructed out of the five letters that still remain.
for (i in seq_len(nrow(best_words))) {
    word_1 <- best_words$letters[[i]]
    
    remaining_letters <- best_letters[!best_letters %in% word_1]
    remaining_words <- best_words %>%
        # This filter call both removes all words that can't be spelled with the remaining letters, and also
        # removes all words that come alphabetically before word_1. This is so we don't have duplicates
        # (since a starter of A/B/C and a different starter of A/C/B are the same thing).
        filter(row_number() > i,
               map_lgl(letters, ~all(.x %in% remaining_letters)))
    
    # If there aren't any words that can be created out of the 10 remaining letters, just skip the rest of
    # this loop iteration and go to the next one.
    if (nrow(remaining_words)==0) {
        next
    }
    
    for (j in seq_len(nrow(remaining_words))) {
        word_2 <- remaining_words$letters[[j]]
        
        still_remaining_letters <- remaining_letters[!remaining_letters %in% word_2]
        still_remaining_words <- remaining_words %>%
            filter(row_number() > j,
                   map_lgl(letters, ~all(.x %in% still_remaining_letters)))
        
        if (nrow(still_remaining_words)==0) {
            next
        }
        
        # If there are any words that can be made out of the five letters that still remain, add them to the
        # opener_combinations table.
        for (k in seq_len(nrow(still_remaining_words))) {
            word_3 <- still_remaining_words$letters[[k]]
            
            opener_combinations <- tibble(word_1=list(word_1),
                                          word_2=list(word_2),
                                          word_3=list(word_3)) %>%
                bind_rows(opener_combinations, .)
        }
    }
}
```

# Cutting the list down

Huh, there are way more valid three word combinations that have all 15 of the most common letters than I thought: 27,154. These openers aren't all equally useful, though, because Wordle is about more than just identifying letters, it's about putting them in the right places. So if all 27,154 of these words are equally good at covering letters, we can rank them by how often they put letters in the right places.

We'll need to find out how often those best letters are in each position, but first, let's tidy up `opener_combinations` a bit.

```{r opener_cleanup}
cleaned_openers <- opener_combinations %>%
    mutate(opener_id=row_number(), .before="word_1") %>%
    pivot_longer(cols=starts_with("word"),
                 names_to="word_num",
                 values_to="letters") %>%
    unnest(letters) %>%
    group_by(opener_id, word_num) %>%
    mutate(letter_position=row_number()) %>%
    ungroup()
```

This pivot/unnest combination has transformed this:

```{r opener_combinations_example, echo=FALSE}
opener_combinations
```

into this:

```{r cleaned_openers_example, echo=FALSE}
cleaned_openers
```

Now, with a tidy dataset to work with, how often does each of the top-15 best letters show up in each position?

```{r best_letters_positions}
best_letters_positions <- split_words %>%
    unnest(letters) %>%
    group_by(words) %>%
    mutate(letter_position=row_number()) %>%
    ungroup() %>%
    group_by(letters, letter_position) %>%
    summarize(position_count=n(), .groups="drop") %>%
    filter(letters %in% best_letters)

best_letters_positions
```

```{r best_letters_positions_viz}
ggplot(best_letters_positions) +
    geom_col(aes(x=letter_position, y=position_count)) +
    facet_wrap(~letters, nrow=3)
```


We have everything we need. Let's grab our cleaned dataframe of opening word combinations, join to the above table with data on how often each letter shows up in each position, and rank the 27,154 combinations to find the ones that get the letters in the right positions the most.

```{r best_openers}
best_openers <- cleaned_openers %>%
    left_join(best_letters_positions, by=c("letters", "letter_position")) %>%
    mutate(position_count=if_else(is.na(position_count), 0L, position_count)) %>%
    group_by(opener_id) %>%
    summarize(position_sum=sum(position_count)) %>%
    arrange(desc(position_sum)) %>%
    inner_join(cleaned_openers, by=c("opener_id")) %>%
    group_by(opener_id, position_sum, word_num) %>%
    summarize(words=toupper(paste(letters, collapse="")), .groups="drop") %>%
    group_by(opener_id, position_sum) %>%
    summarize(opener=paste(words, collapse="/"), .groups="drop") %>%
    arrange(desc(position_sum))

best_openers %>%
    filter(position_sum==max(position_sum)) %>%
    kable()
```
All 28 of these openers are equally good at putting letters in the right place. For my money, the one that's easiest to remember is **CANID/MOLES/PURTY**.
